---
title: "assignment_1"
output: html_document
date: "2023-02-16"
---

Push as Rmd

```{r}
library(tidyverse)
```


First create our agents: 
1. Random agent
2. Win-stay-lose-shift agent
3. Strategy-shifting agent (33% win-stay-lose-shift, 66% win-shift-lose-stay)
4. Win-shift-shift, lose-stay-stay agent

1. Random agent
```{r}
#Defining agent function
RandomAgent_f <- function(rate){
  choice <- rbinom(1, 1, rate)
  return(choice)
}

#Running simulation
#Define rate of choosing 0 or 1
rate <- 0.5
#Define number of trials
trials <- 120
#Create empty vector
randomChoice <- rep(NA, trials)

#Run simulation and save output
for (t in seq(trials)) {randomChoice[t] <- RandomAgent_f(rate)}

#Transform output to tibble
dl <- tibble(trial = seq(trials), choice = randomChoice)

#Plot output
pl <- ggplot(dl, aes(trial, choice)) + geom_line() + theme_classic()
#pl
```

2. Win-stay-lose-shift agent
```{r}
#Defining agent function
WSLSAgent_f <- function(prevChoice, Feedback) {
  if (Feedback == 1) { #If feedback = 1 (win), stay
    choice = prevChoice
  }
  else if (Feedback == 0) { #If feedback = 0 (loss), shift
    choice = 1 - prevChoice
  }
  return(choice)
}
```

3. Strategy-shifting agent (33% win-stay-lose-shift, 66% win-shift-lose-stay)
```{r}
#Define agent function
ProbWSLSAgent_f <- function(prevChoice, Feedback, noise){ 
  
  if(rbinom(1,1, noise) == 1){ 
    
    if (Feedback == 0){
      choice = prevChoice
    } else if (Feedback == 1) {
      choice = 1-prevChoice } }
  
  else {
    if (Feedback == 1){
      choice = prevChoice
    } else if (Feedback == 0) {
      choice = 1-prevChoice } }
  
return(choice) }
```

4. Win-shift-shift, lose-stay-stay agent
```{r}
#Define agent function
WSSLSSAgent_f <- function(prevChoice, Feedback){
  for (i in 2:trials){
  if (i %% 2 == 0) { # if even trial number
    if (Feedback == 1){
    win = TRUE  # keep track of current strategy
    choice = 1-prevChoice 
  } else if (Feedback == 0) { 
    win = FALSE  # keep track of current strategy
    choice = prevChoice }
  }
    
  else if (i %% 2 != 0) { # if odd trial number
    if (win == TRUE) { # if the win-shift-shift strategy
    choice = prevChoice
    if (prevChoice == 1){
      choice = 0
  }   else if (prevChoice == 0) {
        choice = 1}
    
    if (win == FALSE) { # if the lose-stay-stay strategy
    choice = prevChoice
    if (prevChoice == 1){
      choice = 1
  }   else if (prevChoice == 0) {
        choice = 0}}
    }
    }
        }
  return(choice) }
```


Then simulate the games: 
1. Random agent vs strategy-shifting agent (33% win-stay-lose-shift, 67% win-shift-lose-stay)
2. WSLS vs strategy-shifting agent (33% win-stay-lose-shift, 66% win-shift-lose-stay)
3. Random agent vs win-shift-shift, lose-stay-stay agent
4. WSLS vs win-shift-shift, lose-stay-stay agent


1. Random agent vs strategy-shifting agent
```{r}
#Define rate
rate <- 0.5

#Define number of trials
trials = 120

#Define noise (the probability of using strategy win-shift-lose-stay)
noise <- 0.67

#Define empty vectors for dealer and player
Self <- rep(NA, trials) #player
Other <- rep(NA, trials) #dealer

#Define random first choice for player
Self[1] <- RandomAgent_f(0.5)

#Run simulation
for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}
for (i in 2:trials){
  if (Self[i-1] == Other[i-1]){
    Feedback = 1
  } else {Feedback = 0}
  Self[i] <- ProbWSLSAgent_f(Self[i-1], Feedback, noise)
}

#Visualizing 
#The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

#The performance
df <- df %>% mutate(
cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback), cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))
```

For ten participants
```{r}
library(reshape2)

#Define rate
rate <- 0.5

# Define the number of trials and the number of repetitions
trials <- 100
repetitions <- 10

#Define noise (the probability of using strategy win-shift-lose-stay)
noise <- 0.67

#Define empty vectors for dealer and player
Self <- rep(NA, trials) #player
Other <- rep(NA, trials) #dealer

# Create empty matrices to store the results
Self_results <- matrix(0, nrow = trials, ncol = repetitions)
Other_results <- matrix(0, nrow = trials, ncol = repetitions)

#Run simulation
# Outer loop for repetitions
for (j in 1:repetitions) {
  #Define random first choice
  Self[1] <- RandomAgent_f(0.5) #player
  for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}
  for (i in 2:trials){
  if (Self[i-1] == Other[i-1]){
    Feedback = 1
  } else {Feedback = 0}
  Self[i] <- ProbWSLSAgent_f(Self[i-1], Feedback, noise)
  }
  # Store the results in the matrices
  Self_results[, j] <- Self
  Other_results[, j] <- Other
}

#Transform matrices to long (stacked) format
Self_long <- melt(Self_results)
Other_long <- melt(Other_results)

#Merge matrices
merged <- cbind(Self_long, Other_long[,3])

#Rename dataframe variables
names(merged) <- c("Trial", "Participant", "Self_choice", "Other_choice")

#Add feedback
merged$Feedback <- as.numeric(merged$Self_choice == merged$Other_choice)

#Add cumulative performance
merged <- merged %>% 
  mutate(cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
         cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )

#Plotting with dealer/player colors
#ggplot(merged) + theme_classic() + geom_line(color="red", aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(color="blue", aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Participant")

#Plotting according to dyad
ggplot(merged) + theme_classic() + geom_line(aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Dyad")

```


2. WSLS vs strategy-shifting agent
```{r}
#Define rate
rate <- 0.5

#Define number of trials
trials = 120

#Define noise (the probability of using strategy win-shift-lose-stay)
noise <- 0.67

#Define empty vectors
Self <- rep(NA, trials)
Other <- rep(NA, trials)

#Define random first choice for both 
Self[1] <- RandomAgent_f(0.5) #player
Other[1] <- RandomAgent_f(0.5) #dealer

#Run simulation
for (i in 2:trials){
  if(Self[i-1] == Other[i-1]){
    Feedback = 1
  } else {Feedback = 0}
  Self[i] <- ProbWSLSAgent_f(Self[i-1], Feedback, noise)
  Other[i] <- WSLSAgent_f(Other[i-1], Feedback)
}

#Visualizing 
#The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

#The performance
df <- df %>% mutate(
cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback), cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))
```

For ten participants
```{r}
library(reshape2)

#Define rate
rate <- 0.5

#Define noise (the probability of using strategy win-shift-lose-stay)
noise <- 0.67

# Define the number of trials and the number of repetitions
trials <- 100
repetitions <- 10

#Define empty vectors
Self <- rep(NA, trials)
Other <- rep(NA, trials)

# Create empty matrices to store the results
Self_results <- matrix(0, nrow = trials, ncol = repetitions)
Other_results <- matrix(0, nrow = trials, ncol = repetitions)

#Define random first choice for both 
Self[1] <- RandomAgent_f(0.5) #player
Other[1] <- RandomAgent_f(0.5) #dealer


#Run simulation
# Outer loop for repetitions
for (j in 1:repetitions) {
  #Define random first choice
  Self[1] <- RandomAgent_f(0.5) #player
  Other[1] <- RandomAgent_f(0.5) #dealer
  for (i in 2:trials){
  if(Self[i-1] == Other[i-1]){
    Feedback = 1
  } else {Feedback = 0}
  Self[i] <- ProbWSLSAgent_f(Self[i-1], Feedback, noise)
  Other[i] <- WSLSAgent_f(Other[i-1], Feedback)
}
  # Store the results in the matrices
  Self_results[, j] <- Self
  Other_results[, j] <- Other
}

#Transform matrices to long (stacked) format
Self_long <- melt(Self_results)
Other_long <- melt(Other_results)

#Merge matrices
merged <- cbind(Self_long, Other_long[,3])

#Rename dataframe variables
names(merged) <- c("Trial", "Participant", "Self_choice", "Other_choice")

#Add feedback
merged$Feedback <- as.numeric(merged$Self_choice == merged$Other_choice)

#Add cumulative performance
merged <- merged %>% 
  mutate(cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
         cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )

#Plotting with dealer/player colors
#ggplot(merged) + theme_classic() + geom_line(color="red", aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(color="blue", aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Participant")

#Plotting according to dyad
ggplot(merged) + theme_classic() + geom_line(aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Dyad")
```


3. Random agent vs win-shift-shift, lose-stay-stay agent
```{r}
#Define rate
rate <- 0.5

#Define number of trials
trials = 120

#Define empty vectors
Self <- rep(NA, trials)
Other <- rep(NA, trials)

#Define random first choice for player
Self[1] <- RandomAgent_f(0.5) #player

#Run simulation
for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}
for (i in 2:trials){
  if (Self[i-1] == Other[i-1]){
    Feedback = 1
  } else {Feedback = 0}
  Self[i] <- WSSLSSAgent_f(Self[i-1], Feedback)
}

#Visualizing 
#The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other)) 
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

#The performance
df <- df %>% 
  mutate(cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
         cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )
df


ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))
```

For ten participants
```{r}
library(reshape2)

#Define rate
rate <- 0.5

# Define the number of trials and the number of repetitions
trials <- 100
repetitions <- 10

#Define empty vectors
Self <- rep(NA, trials)
Other <- rep(NA, trials)

# Create empty matrices to store the results
Self_results <- matrix(0, nrow = trials, ncol = repetitions)
Other_results <- matrix(0, nrow = trials, ncol = repetitions)

#Define random first choice for both 
Self[1] <- RandomAgent_f(0.5) #player
Other[1] <- RandomAgent_f(0.5) #dealer

#Run simulation
# Outer loop for repetitions
for (j in 1:repetitions) {
  #Define random first choice
  Self[1] <- RandomAgent_f(0.5) #player
  for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}
  for (i in 2:trials){
  if (Self[i-1] == Other[i-1]){
    Feedback = 1
  } else {Feedback = 0}
  Self[i] <- WSSLSSAgent_f(Self[i-1], Feedback)
}
  # Store the results in the matrices
  Self_results[, j] <- Self
  Other_results[, j] <- Other
}

#Transform matrices to long (stacked) format
Self_long <- melt(Self_results)
Other_long <- melt(Other_results)

#Merge matrices
merged <- cbind(Self_long, Other_long[,3])

#Rename dataframe variables
names(merged) <- c("Trial", "Participant", "Self_choice", "Other_choice")

#Add feedback
merged$Feedback <- as.numeric(merged$Self_choice == merged$Other_choice)

#Add cumulative performance
merged <- merged %>% 
  mutate(cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
         cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )

#Plotting with dealer/player colors
#ggplot(merged) + theme_classic() + geom_line(color="red", aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(color="blue", aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Participant")

#Plotting according to dyad
ggplot(merged) + theme_classic() + geom_line(aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Dyad")
```


4. WSLS vs win-shift-shift, lose-stay-stay agent
```{r}
#Define empty vectors
Self <- rep(NA, trials) 
Other <- rep(NA, trials) 

#Define number of trials
trials = 120

#Define random first choice
Self[1] <- RandomAgent_f(0.5) #player
Other[1] <- RandomAgent_f(0.5) #dealer

#Run simulation
for (i in 2:trials){ 
  if (Self[i-1] == Other[i-1]){ Feedback = 1 } else {Feedback = 0}
  Self[i] <- WSSLSSAgent_f(Self[i-1], Feedback) 
  Other[i] <- WSLSAgent_f(Other[i-1], 1-Feedback) }

#Visualizing 
#The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other)) 
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

#The performance
df <- df %>% 
  mutate(cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
         cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )

ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))

```


For ten participants
```{r}
library(reshape2)

# Define the number of trials and the number of repetitions
trials <- 100
repetitions <- 10

# Create empty matrices to store the results
Self_results <- matrix(0, nrow = trials, ncol = repetitions)
Other_results <- matrix(0, nrow = trials, ncol = repetitions)

# Outer loop for repetitions
for (j in 1:repetitions) {
  #Define random first choice
  Self[1] <- RandomAgent_f(0.5) #player
  Other[1] <- RandomAgent_f(0.5) #dealer
  
  # Inner loop for trials
  for (i in 2:trials) {
    if (Self[i-1] == Other[i-1]) {
      Feedback = 1
    } else {
      Feedback = 0
    }
    Self[i] <- WSSLSSAgent_f(Self[i-1], Feedback)
    Other[i] <- WSLSAgent_f(Other[i-1], 1-Feedback)
  }
  
  # Store the results in the matrices
  Self_results[, j] <- Self
  Other_results[, j] <- Other
}

#Transform matrices to long (stacked) format
Self_long <- melt(Self_results)
Other_long <- melt(Other_results)

#Merge matrices
merged <- cbind(Self_long, Other_long[,3])

#Rename dataframe variables
names(merged) <- c("Trial", "Participant", "Self_choice", "Other_choice")

#Add feedback
merged$Feedback <- as.numeric(merged$Self_choice == merged$Other_choice)

#Add cumulative performance
merged <- merged %>% 
  mutate(cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
         cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )

#Plotting with dealer/player colors
#ggplot(merged) + theme_classic() + geom_line(color="red", aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(color="blue", aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Participant")

#Plotting according to dyad
ggplot(merged) + theme_classic() + geom_line(aes(x = Trial, y = cumulativerateSelf, group = Participant, color = as.factor(Participant))) + geom_line(aes(x = Trial, y = cumulativerateOther, group = Participant, color = as.factor(Participant))) + labs(x = "Trial", y = "Choice", color = "Dyad")
```





